---
title: "Act10"
output: html_document
date: "2026-06-22"
---
```{r}
library(maps)
library(osmdata)
library(sf)
library(terra)
library(ggplot2)
library(dplyr)
library(ggnewscale)
```

A3: Clean the data by removing records found outside the reported species range. 
(data is cleaned by manually filtered out blanks and removing records found outside the reported species range in excel)
```{r}
df10 <- read.csv("Trimeresurus schultzei.csv")
```

A4: Go to the DIVA-GIS website and download the Philippine administrative boundaries/areas. The zip file contains multiple shapefiles (adm0–3), representing different boundary levels (0-country; 1-provincial; 2-municipal; 3-barangay).
```{r}
adm <- st_read("PHL_adm1.shp")
```

A5: Get only the provincial level, and subset the provinces that contain only distribution records for your assigned species.
```{r}
pal <- adm[adm$NAME_1 == "Palawan", ]
```

A6: Go back to the DIVA-GIS website. Download the Philippine digital elevation model (“Elevation [country mask]”), then clip it using the provinces you selected in step 5.
```{r}
alt <- rast("PHL_msk_alt.vrt")
print(st_crs(pal) != st_crs(alt))
pal <- st_transform(pal, st_crs(alt))
dem <- crop(alt, pal) |> mask(pal)
```

B1: Now, put all the data together into a map, following the pre-lab lecture. Make sure that data layers (points, polygons, DEMs) are arranged properly. Be creative!
```{r}
names(dem) <- "elevation"
r_df <- as.data.frame(dem, xy = TRUE)
ggplot() +
  geom_tile(data = r_df, aes(x = x, y = y, fill = elevation)) +
  scale_fill_gradientn(
    colors = c("#2b83ba", "#abdda4", "#ffffbf", "#fdae61", "#d7191c"),
    name = "Elevation (m)") +
  new_scale_fill() +
  geom_point(data = df10,
             aes(x = decimalLongitude, y = decimalLatitude),
             shape = 21,
             fill = "black",
             color = "green",
             size = 3,
             stroke = 1.2) +
  coord_equal() +
  labs(title = "Palawan Province",
       x = "Longitude",
       y = "Latitude") +
  theme_bw()
```
