---
title: "Act09"
output: html_document
date: "2026-06-17"
---

A2: 
Read in the frogs’ data as a dataframe.
```{r}
df09 <- read.csv("Meneses_et.al_2026_Platymantis.csv")
```

A3:
Subset your data to retain only the “Species” column and morphological features (e.g., SVL, HL, SNL, etc.) from columns 10 through the last one (i.e., 10:X).
```{r}
df09 <- df09[, c(1, 10:34)]
```

A4: 
Check the class and structure of the subset data. How many morphological features are present in the dataset?
```{r}
class(df09)
str(df09)
```

A5: 
Ensure the dataframe has no NAs.
```{r}
good <- complete.cases(df09)
df09 <- df09[good, ]
```

A6: 
Scale the data using only the numeric columns, then compute the pairwise correlations. Do not further subset the data.
```{r}
df09_scaled <- scale(df09[, 2:26])
cor <- cor(df09_scaled)
round(cor, 2)
```

A7: 
Use the correlation matrix to calculate the eigenvectors and the proportion of variance (expressed in percentage). How many principal components were produced? What is the proportion of variance explained by components 1–4? 
```{r}
eig <- eigen(cor)
prop_var <- eig$values / sum(eig$values)
prop_var_per <- (eig$values / sum(eig$values)) * 100
prop_var_per
length(prop_var_per)
prop_var_per[1:4]
```

A8: 
Now, create a 1x2 plot showing the screeplot and the PCA biplot. You can use either base R or ggplot2. You can be creative here by coloring the point shapes per species and adding a legend. Consider the high data-ink ratio you learned from the lecture.
```{r}
par(mfrow = c(1,2))
barplot(prop_var,
        names.arg = paste0("PC", 1:25),
        ylim = c(0, 1),
        col = "orange",
        main = "Scree Plot",
        ylab = "Proportion of Variance"
        )
eig_vec <- eig$vectors
scores <- df09_scaled %*% eig_vec
head(scores)
plot(scores[,1], scores[,2],
     xlab = "PC1",
     ylab = "PC2",
     main = "PCA Score Plot (PC1 vs PC2)",
     pch = 19,
     col = as.factor(df09$Species)
)
legend("topright",
       legend = levels(as.factor(df09$Species)),
       col = 1:2,
       pch = 19)
```

B1: 
Examine the variance explained by principal components using the Broken-stick method and the PCAtest package. What is the difference between these two methods for assessing the importance of principal components?
```{r}
broken_stick <- function(p){
  sapply(1:p, function(k){
    sum(1 / (k:p)) / p
  })
}
bs <- broken_stick(25)
bs
comparison <- data.frame(
  PC = paste0("PC", 1:25),
  Observed = prop_var,
  BrokenStick = bs
)
comparison
```

B2: 
Create a barplot showing the results of Broken-stick analysis and perform a significance test for principal components using PCAtest, following the instructions at https://arleyc.github.io/PCAtest/.
```{r}
par(mfrow = c(1,1))
plot(prop_var,
     type = "b",
     pch = 19,
     ylim = c(0, max(prop_var)),
     xlab = "Principal Component",
     ylab = "Proportion of Variance",
     main = "Broken-Stick Analysis")
lines(bs,
      type = "b",
      pch = 17,
      col = "red")
legend("topright",
       legend = c("Observed", "Broken Stick"),
       pch = c(19, 17),
       col = c("black", "red"))
library(PCAtest)
result <- PCAtest(df09_scaled , 1000, 1000, 0.5, varcorr = FALSE, counter = FALSE, plot = TRUE)
```
