---
title: "Act07"
output: html_document
date: "2026-06-08"
---

A2: 
Read in the morphology data as a dataframe.
```{r}
df07 <- read.csv("morphology_data.csv")
```

A3: 
Remove the sex column.
```{r}
df07 <- df07[, -2]
df07 <- df07[, -1]
```

A4: 
Ensure that the dataframe has no NAs.
```{r}
good <- complete.cases(df07)
df07 <- df07[good, ]
```

A5: 
Prior to performing K-means, standardize all values using the “scale” function.
```{r}
df07_scaled <- scale(df07)
```

A6: 
Perform K-means with 3 cluster centers using 25 initial random centroids.
```{r}
set.seed(123)
kmeans_result <- kmeans(df07_scaled, centers = 3, nstart = 25)
```

A7: 
Show the cluster assignment and a summary of results by simply printing them to the console.
```{r}
head(kmeans_result$cluster)
kmeans_result
```

A8: 
Create a scatter plot with points colored by the assigned cluster.
```{r}
plot(
  df07_scaled[, 1:2],
  col = kmeans_result$cluster,
  pch = 19,
  xlab = "SVL (scaled)",
  ylab = "TD (scaled)",
  main = "K-means Clustering"
)
```

B1: 
Now, create a loop function to perform K-means with cluster centers from 2 to 10.
```{r}
wss <- 0
for (k in 2:10){
  set.seed(123) 
  kmeans210 <- kmeans(df07_scaled, centers = k, nstart = 25)
  wss[k] <- kmeans210$tot.withinss
}
```

B2: 
From the output of your loop function, create an elbow curve plot to determine the optimal K.
```{r}
plot(2:10, 
     wss[2:10], 
     type = "b", 
     pch = 19,
     xlab = "Number of Clusters (K)", 
     ylab = "WSS", 
     main = "Elbow Curve"
     )
library(ggplot2)
library(factoextra)
fviz_nbclust(
  df07_scaled,
  kmeans,
  method = "wss",
  k.max = 10
)
```

B3: 
Recreate steps 6–8 above using the optimal K determined by the elbow curve.
```{r}
set.seed(123) 
kmeans2 <- kmeans(df07_scaled, centers = 2, nstart = 25)
kmeans2$cluster
kmeans2
plot(
  df07_scaled[, 1:2],
  col = kmeans2$cluster,
  pch = 19,
  xlab = "SVL (scaled)",
  ylab = "TD (scaled)",
  main = "K-means Clustering (K=2)"
)
```
