---
title: "Act11"
output: html_document
date: "2026-07-06"
---
A2: Read in the species data and remove unnecessary rows with NAs.
```{r}
df11 <- read.csv("herpetozoa-033-095-s001.csv")
good <- complete.cases(df11)
df11 <- df11[good, ]
```

A3: Read in the elevation data (ph_strm_geo.tif).
```{r}
library(terra)
elev <- rast("wc2.1_10m_elev.tif")
```

A4: From the species data, subset the longitude and latitude coordinates. Then convert the coordinates to a spatial vector using the “vect” function from the “terra” package.
```{r}
coord <- vect(df11, geom = c("Long", "Lat"))
```

A5: Using the converted coordinates, extract elevation data for each point with the “extract” function in the “terra” package. The extracted data will have two columns: “ID” and “Band_1.” The “Band_1” column contains the elevation.
```{r}
extractElev <- extract(elev, coord)
```

A6: Subset the species data to retain only the species richness column (SPPTotal). Then combine this column with the extracted elevation. Rename “Band_1” to “elevation.”
```{r}
df11SPP <- data.frame(
  SPPTotal = df11$SPPTotal,
  elevation = extractElev[, 2]
)
```

A7: Now, perform simple linear regression using the code from the pre-lab lecture, including a plot. What can you infer about the relationship between species and elevation?
```{r}
x <- df11SPP$elevation
y <- df11SPP$SPPTotal
xbar <- mean(x)
ybar <- mean(y)
b <- sum((x - xbar) * (y - ybar)) /
     sum((x - xbar)^2)
a <- ybar - b * xbar
cat("Regression Equation:\n")
cat("y =", round(a,3), "+", round(b,3), "x\n")
yhat <- a + b * x
plot(x, y,
     pch = 19,
     col = "orange",
     xlab = "Elevation (m)",
     ylab = "SPPTotal",
     main = "Simple Linear Regression")
abline(a, b,
       col = "blue",
       lwd = 2)
points(x, yhat,
       pch = 17,      # Triangle
       col = "green",
       cex = 1.2)
segments(x0 = x, y0 = y,
         x1 = x, y1 = yhat,
         lty = 2,
         col = "red")
legend("topright",
       legend = c("Observed", "Predicted", "Regression Line"),
       pch = c(19, 17, NA),
       lty = c(NA, NA, 1),
       col = c("orange", "green", "blue"),
       lwd = c(NA, NA, 2),
       bty = "n",
       cex = 0.8)
```

B1: Read in both annual precipitation and temperature data and extract the data following the instructions from A4–5.
```{r}
preci <- rast("annual_precipitation.tif")
temp <- rast("annual_temperature.tif")
extractpreci <- extract(preci, coord)
extracttemp <- extract(temp, coord)
```

B2: Combine the extracted data with the species and elevation data into a single dataframe.
```{r}
df11Ult <- data.frame(
  SPPTotal = df11$SPPTotal,
  elevation = extractElev[, 2],
  precipitation = extractpreci[, 2],
  temperature = extracttemp[, 2]
)
```

B3: Perform multiple linear regression using the “lm” function.
```{r}
x <- df11Ult$elevation
x1 <- df11Ult$precipitation
x2 <- df11Ult$temperature
y <- df11Ult$SPPTotal
model <- lm(y ~ x + x1 + x2)
coef(model)
a <- coef(model)[1]
b <- coef(model)[2]
c <- coef(model)[3]
d <- coef(model)[4]
cat("Regression Equation:\n")
cat("y =", round(a, 3), "+", round(b, 3), "x +", round(c, 3), "x1 +", round(d, 3), "x2\n\n")
yhat <- predict(model)
residuals <- residuals(model)
summary(model)$r.squared
results <- data.frame(
  elevation = x, 
  precipitation = x1,
  temperature = x2, 
  Observed = y,
  Predicted = round(yhat,2),
  Residual = round(residuals,2)
)
print(results)
summary(model)
```

B4: Which relationship is statistically significant?
The relationship of species diversity with 3 conditions: elevation, precipitation, and temperature, are all statistically significant. 
